---
title: "The Harry Potter Colour Palettes"
author:
  - "Alejandro Jiménez Rico"
date: "`r Sys.Date()`"
output:
  md_document:
    variant: markdown_github
    toc: true
    toc_depth: 1
---

<style>
img {
    max-width: 100%;
    max-height: 100%;
}
</style>

# tl;dr

Use the color scales in this package to make plots that make use of the palette extracted from the *Harry Potter* movie franchise.

Install **hp** like any R package:

```{r}
library(devtools)
devtools::install_github("aljrico/harrypotter", force = TRUE)
library(harrypotter)
```

For base plots, use the `hp()` function to generate a palette:

```{r setup, include=FALSE}
library(harrypotter)
knitr::opts_chunk$set(echo = TRUE, fig.retina=2, fig.width=7, fig.height=5)
```

```{r tldr_base, message=FALSE}
x <- y <- seq(-8*pi, 8*pi, len = 40)
r <- sqrt(outer(x^2, y^2, "+"))
filled.contour(cos(r^2)*exp(-r/(2*pi)), 
               axes=FALSE,
               color.palette=hp,
               asp=1)
```

For ggplot, use `scale_colour_hp()` and `scale_fill_hp()`:

```{r, tldr_ggplot, message=FALSE}
library(ggplot2)
ggplot(data.frame(x = rnorm(10000), y = rnorm(10000)), aes(x = x, y = y)) +
  geom_hex() + coord_fixed() +
  scale_fill_hp(movie = "hufflepuff") + theme_bw()
```

# Introduction

The [**harrypotter**](http:://github.com/aljrico/harrypotter) package
brings to R colour scales created by [Alejandro Jimenez Rico](https://github.com/aljrico) from the [**Harry Potter** film series](https://en.wikipedia.org/wiki/Harry_Potter_(film_series)) 

# The Colour Scales

The package contains eight color scales: One for each movie of the franchise.

```{r for_repeat, include=FALSE}
n_col <- 128

img <- function(obj, nam) {
  image(1:length(obj), 1, as.matrix(1:length(obj)), col=obj, 
        main = nam, ylab = "", xaxt = "n", yaxt = "n",  bty = "n")
}
```

```{r begin, message=FALSE, include=FALSE}
library(harrypotter)
library(scales)
library(colorspace)
library(dichromat)
```

### Movies
```{r show_scales, echo=FALSE,fig.height=3.575}
par(mfrow=c(8, 1), mar=rep(1, 4))
img(rev(hp(n_col, movie = 1)), 1)
img(rev(hp(n_col, movie = 2)), 2)
img(rev(hp(n_col, movie = 3)), 3)
img(rev(hp(n_col, movie = 4)), 4)
img(rev(hp(n_col, movie = 5)), 5)
img(rev(hp(n_col, movie = 6)), 6)
img(rev(hp(n_col, movie = 7)), 7)
img(rev(hp(n_col, movie = 8)), 8)
```

### Houses
```{r show_scales2, echo=FALSE,fig.height=4}
par(mfrow=c(4, 1), mar=rep(1, 4))
img(rev(hp(n_col, movie = "slytherin")), "Slytherin")
img(rev(hp(n_col, movie = "gryffindor")), "Gryffindor")
img(rev(hp(n_col, movie = "ravenclaw")), "Ravenclaw")
img(rev(hp(n_col, movie = "hufflepuff")), "Hufflepuff")
```

# Usage

The `hp()` function produces the Harry Potter color scale.  You can choose
the other color scale options using the `movie` parameter.

The package also contains color scale functions for **ggplot**
plots: `scale_color_hp()` and `scale_fill_hp()`.  As with `hp()`, you can use the other scales with the `movie` argument
in the `ggplot` scales.  Here the scale from the 7th is used for a cloropleth
map of U.S. unemployment:

```{r, ggplot2}
unemp <- read.csv("http://datasets.flowingdata.com/unemployment09.csv",
                  header = FALSE, stringsAsFactors = FALSE)
names(unemp) <- c("id", "state_fips", "county_fips", "name", "year",
                  "?", "?", "?", "rate")
unemp$county <- tolower(gsub(" County, [A-Z]{2}", "", unemp$name))
unemp$county <- gsub("^(.*) parish, ..$","\\1", unemp$county)
unemp$state <- gsub("^.*([A-Z]{2}).*$", "\\1", unemp$name)

county_df <- map_data("county", projection = "albers", parameters = c(39, 45))
names(county_df) <- c("long", "lat", "group", "order", "state_name", "county")
county_df$state <- state.abb[match(county_df$state_name, tolower(state.name))]
county_df$state_name <- NULL

state_df <- map_data("state", projection = "albers", parameters = c(39, 45))

choropleth <- merge(county_df, unemp, by = c("state", "county"))
choropleth <- choropleth[order(choropleth$order), ]

ggplot(choropleth, aes(long, lat, group = group)) +
  geom_polygon(aes(fill = rate), colour = alpha("white", 1 / 2), size = 0.2) +
  geom_polygon(data = state_df, colour = "white", fill = NA) +
  coord_fixed() +
  theme_minimal() +
  ggtitle("US unemployment rate by county") +
  theme(axis.line = element_blank(), axis.text = element_blank(),
        axis.ticks = element_blank(), axis.title = element_blank()) +
  scale_fill_hp(movie="gryffindor")

ggplot(choropleth, aes(long, lat, group = group)) +
  geom_polygon(aes(fill = rate), colour = alpha("white", 1 / 2), size = 0.2) +
  geom_polygon(data = state_df, colour = "white", fill = NA) +
  coord_fixed() +
  theme_minimal() +
  ggtitle("US unemployment rate by county") +
  theme(axis.line = element_blank(), axis.text = element_blank(),
        axis.ticks = element_blank(), axis.title = element_blank()) +
  scale_fill_hp(movie="slytherin")

ggplot(choropleth, aes(long, lat, group = group)) +
  geom_polygon(aes(fill = rate), colour = alpha("white", 1 / 2), size = 0.2) +
  geom_polygon(data = state_df, colour = "white", fill = NA) +
  coord_fixed() +
  theme_minimal() +
  ggtitle("US unemployment rate by county") +
  theme(axis.line = element_blank(), axis.text = element_blank(),
        axis.ticks = element_blank(), axis.title = element_blank()) +
  scale_fill_hp(movie="ravenclaw")

ggplot(choropleth, aes(long, lat, group = group)) +
  geom_polygon(aes(fill = rate), colour = alpha("white", 1 / 2), size = 0.2) +
  geom_polygon(data = state_df, colour = "white", fill = NA) +
  coord_fixed() +
  theme_minimal() +
  ggtitle("US unemployment rate by county") +
  theme(axis.line = element_blank(), axis.text = element_blank(),
        axis.ticks = element_blank(), axis.title = element_blank()) +
  scale_fill_hp(movie="hufflepuff")
```

The ggplot functions also can be used for discrete scales with the argument
`discrete=TRUE`.

```{r discrete}
p <- ggplot(mtcars, aes(wt, mpg))
p + geom_point(size=4, aes(colour = factor(carb))) +
    scale_color_hp(discrete=TRUE, movie = 1) +
    theme_bw()
```

